### 2024-09-27 paper/paper.qmd
update paper

---
title: "Who's Homeless in Toronto: Major Trends in Single Adults Aged 25 to 44 Becoming Homeless from 2018 to 2024"
author: 
  - Xizi Sun
thanks: "A GitHub Repository containing all data, R code, and other files used in this investigation is located here: https://github.com/LilianS77/Toronto_shelter_research.git"
date: today
date-format: long
abstract: "This study examines homelessness trends in Toronto from 2018 to 2024, focusing on demographic shifts and socio-economic factors. Findings show a sharp rise in homelessness among single adults aged 25-44, largely due to post-pandemic unemployment. The research provides insights into how Toronto can better allocate resources to reduce homelessness by targeting vulnerable groups. Stronger public policies focused on expanding support services and employment opportunities for at-risk populations are recommended."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(opendatatoronto)
library(tidyverse)
library(knitr)
library(ggplot2)
library(dplyr)
library(lubridate)
```

# Introduction {#sec-introduction}

In recent years, the growing number of homeless people in Toronto has raised concerns and alarms, generating a high level of public concern and calls for action. This paper uses dataset to analyze age group trends and demographic dynamics of Toronto's homeless population, using the “Toronto Shelter System Flows” dataset [@torontosheltersystem] that tracks homelessness from 2018 to 2024. By focusing on how different age groups and populations are affected, the study delves into the rising rates of homelessness among single adults aged 25-44. It also discusses the post-pandemic rise in unemployment as a socio-economic factor contributing to these trends [@jones2020canadianlabourmarket].

Despite the growing awareness of Toronto's homelessness crisis, most existing research has not adequately discussed the changing composition of the homeless population. This study aims to fill this gap by analyzing these demographic changes in detail and linking them to broader societal trends. By identifying the groups most at risk, the study provides important insights into how the city of Toronto can better allocate resources to prevent homelessness.

In this report, @sec-data-tools describes the data tools I used to analyze the data, and @sec-data-source-and-measurement explains the context and measurements used with this dataset. @sec-data-selection explains the reasons for not choosing another database. @sec-Results presents and describes the demographic distribution and trends of homelessness in graphs and charts. @sec-Discussion analyzes the information in the graphs in the context of the community and describes the limitations of the study.

# Data {#sec-data}

## Data Tools {#sec-data-tools}
All data analysis and visualizations were conducted using R [@citeR]. Data was obtained from the City of Toronto's Open Data Catalogue [@opendata] and included the Toronto Shelter System Flow dataset [@torontosheltersystem]. Visualizations were created using the ggplot2 package [@ggplot2], with data manipulation carried out via dplyr [@dplyr] and tidyr [@tidyr]. Date handling was simplified with the lubridate package [@lubridate], and all R code was formatted using the styler package [@styler]. Guidance on storytelling with data was drawn from Telling Stories with Data [@tellingstorieswithdata].

## Data Source And Measurement {#sec-data-source-and-measurement}
The dataset Toronto Shelter System Flow[@torontosheltersystem] used in this study is sourced from the Toronto Open Data Platform[@opendata]. It measures and tracks homelessness in Toronto by capturing information through a Shelter Management Information System (SMIS) that covers various services such as shelters, 24-hour respite sites, and warming centers funded by the City of Toronto. The data spans from January 2018 to September 2024 and includes individuals who have accessed any of these services in the past three months. The dataset updated monthly and published on the 15th of each month, provides monthly records on the movement of people into and out of these services, offering insight into the demographics, trends, and dynamics of homelessness in Toronto.

The table [@tbl-cleaned-data] shows the variables after data cleaning, showing 6 of the 604 observations. Due to space constraints we are only showing 2 age groups in this table. The variable “date.mmm.yyy.” goes by month from 2018 to August 2024. The data covers different age groups: under 16, 16-24, 25-34, 35-44, 45-54, 55-64, 65 and over, and types of population: refugees, single adults, families, youth, and chronically ill. It tracks inflow and outflow: returned from shelter and returned from housing.

```{r}
#| label: tbl-cleaned-data
#| tbl-cap: Sample of cleaned lead data 
#| echo: false

# Load the dataset
cleaned_data <- read_csv(here::here("data", "analysis_data", "analysis_data.csv"), show_col_types = FALSE)

# Display the first 6 rows of selected columns and match with column names
head(cleaned_data[, c("date.mmm.yy.", "population_group", "returned_from_housing", "returned_to_shelter", "ageunder16", "age16.24")], 6) |>
  kable(
    col.names = c("Date", "Population Group", "Returned from Housing", "Returned to Shelter", "Age under 16", "Age 16-24"),
    booktabs = TRUE
  )

```

## Data Selection {#sec-data-selection}

There are many databases on homelessness in Toronto, such as About Hostel Services: Homeless Shelter Locations[@toronto2019hostelservices], but it doesn't distinguish well between population types, and we can't tell from it where homeless people come from. The data [@torontosheltersystem] I chose has very complete age groupings and also divides the people who seek help from shelters like they return from housing or return to shelter. better still we can know more information about the homeless from this dataset such as whether they are refugees, single or have a family so that we can combine it with the literature to have a better discussion about the Homeless trends and causes of homelessness in Toronto.

# Result {#sec-result}

```{r}
#| echo: false
#| warning: false
#| label: fig-total-homeless-trend-by-quarter
#| fig-cap: "Total Homeless Population Trend by Quarter"

# Generated by ChatGPT

# Define the age group columns
age_columns <- c("ageunder16", "age16.24", "age25.34", "age35.44", "age45.54", "age55.64", "age65over")

# Convert the date column to a proper date format, extract year and quarter, and filter out 2024 Q3
total_homeless_quarter_trend <- cleaned_data %>%
  mutate(
    date = as.Date(paste0("01-", date.mmm.yy.), format = "%d-%b-%y"),
    year = year(date),
    quarter = quarter(date)
  ) %>%
  filter(!(year == 2024 & quarter == 3)) %>%
  group_by(year, quarter) %>%
  summarise(total_homeless = sum(across(all_of(age_columns)), na.rm = TRUE), .groups = "drop") %>%
  mutate(period = paste0(year, " Q", quarter)) # Create a combined year-quarter label

# Create a line plot to show the total homeless population trend by quarter (excluding 2024 Q3)
ggplot(total_homeless_quarter_trend, aes(x = period, y = total_homeless, group = 1)) +
  geom_line(color = "blue") +
  geom_point(color = "blue") +
  labs(
    x = "Year-Quarter", y = "Total Homeless Population"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
@fig-total-homeless-trend-by-quarter shows the total homeless population trend from 2018 Q1 to 2024 Q2. The population rises until 2019, then sharply declines in 2020, with the lowest point falling below 80,000. From 2021 onward, the population steadily increases, reaching its highest point in 2024 Q2, exceeding 120,000.
```{r}
#| echo: false
#| warning: false
#| label: fig-returned-to-shelter
#| fig-cap: "Returned to Shelter Across Population Groups"

# Load dplyr to ensure filter works
library(dplyr)

#Generated by ChatGPT

# Assuming 'cleaned_data' or 'data' is your dataframe
ggplot(
  cleaned_data %>% dplyr::filter(population_group != "All Population"),
  aes(x = population_group, y = returned_to_shelter, fill = population_group)
) +
  geom_bar(stat = "identity") +
  labs(
  x = "Population Group", y = "Returned to Shelter"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
@fig-returned-to-shelter describes a significantly higher number of non-refugees and single adults returning to shelter compared to other groups. Approximately 30,000 non-refugees returned to shelter, the highest value in the chart. Single adults are slightly below 30,000, but still make up a large percentage of the population returning to shelters. Refugee, family, chronically ill, indigenous and youth groups returned to shelters in significantly lower numbers.
```{r}
#| echo: false
#| warning: false
#| label: fig-returned-from-housing
#| fig-cap: "Returned from Housing Across Population Groups"

#Generated by ChatGPT

# Ensure you are referencing the correct dataset, like 'cleaned_data'
ggplot(
  cleaned_data %>% dplyr::filter(population_group != "All Population"),
  aes(x = population_group, y = returned_from_housing, fill = population_group)
) +
  geom_bar(stat = "identity") +
  labs(
    x = "Population Group", y = "Returned from Housing"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
@fig-returned-from-housing shows shelter returns among different demographic groups, and similar to returns to shelters, non-refugees had the highest number of returns from shelter, just over 4,000. The number of single adults is also high at just under 4,000. Chronically ill and family populations are moderately represented, with both returning from shelters in the range of 1,000 to 1,500 people. The indigenous and youth groups have the lowest rates of return from housing, with just over 500 indigenous people and just over 1,000 youth.

```{r}
#| echo: false
#| warning: false
#| label: fig-age-group-distribution-single-adult-pie
#| fig-cap: "Age Group Distribution in Single Adult Population"

#generated by ChatGPT

# Define the age group columns
age_columns <- c("ageunder16", "age16.24", "age25.34", "age35.44", "age45.54", "age55.64", "age65over")

# Filter for Single Adult Population, adjust this filter condition based on your actual dataset structure
single_adult_population <- cleaned_data %>%
  filter(population_group == "Single Adult") %>%
  pivot_longer(cols = all_of(age_columns), names_to = "Age Group", values_to = "Count") %>%
  group_by(`Age Group`) %>%
  summarise(total_population = sum(Count, na.rm = TRUE), .groups = "drop")

# Calculate percentages
single_adult_population <- single_adult_population %>%
  mutate(percentage = total_population / sum(total_population) * 100,
         label = paste0(round(percentage, 1), "%"))  # Create a label with percentage

# Plot the pie chart
ggplot(single_adult_population, aes(x = "", y = percentage, fill = `Age Group`)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = label), position = position_stack(vjust = 0.5)) +  # Add percentage labels
  theme_void() +  # Remove unnecessary axes
  theme(legend.position = "right") +
  scale_fill_brewer(palette = "Set3")  # Optional: Add color palette
```
@fig-age-group-distribution-single-adult-pie shows the distribution of age groups within the current single adult homeless population, highlighting the predominance of the 25.34 and 35.44 year old age groups, which account for approximately 25.2% and 25.8% of the total population respectively. In contrast, the 16.24 year old age group is relatively underrepresented at 8.5%. 65 years and over are not represented in any of the single adult populations, accounting for 0%. The under-16s are minors and therefore account for 0%. This distribution suggests that middle-aged people form the bulk of the population, while the young and the old are virtually absent.
```{r}
#| echo: false
#| warning: false
#| label: fig-age-group-proportion-trend-yearly
#| fig-cap: "Trends in Age Group Proportions Over Time (Yearly)"

#developed by ChatGPT

# Define the age group columns
age_columns <- c("ageunder16", "age16.24", "age25.34", "age35.44", "age45.54", "age55.64", "age65over")

# Convert the date column to a proper date format, extract year, and calculate the total population and proportions for each age group
age_group_proportion_trend_yearly <- cleaned_data %>%
  mutate(
    date = as.Date(paste0("01-", date.mmm.yy.), format = "%d-%b-%y"),
    year = year(date)
  ) %>%
  pivot_longer(cols = all_of(age_columns), names_to = "Age Group", values_to = "Count") %>%
  group_by(year, `Age Group`) %>%
  summarise(total_population = sum(Count, na.rm = TRUE), .groups = "drop") %>%
  group_by(year) %>%
  mutate(Proportion = total_population / sum(total_population)) %>%
  ungroup()

# Create a line plot to show the age group proportion trend over time (yearly)
ggplot(age_group_proportion_trend_yearly, aes(x = factor(year), y = Proportion, color = `Age Group`, group = `Age Group`)) +
  geom_line(size = 1.2) +
  geom_point() +
  scale_y_continuous(labels = scales::percent) +  # Format y-axis as percentages
  labs(
    x = "Year", y = "Proportion of Population"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```
@fig-age-group-proportion-trend-yearly shows several important trends in the homeless population by age group. The 25-34 and 35-44 age groups have been consistently higher than other age groups since 2018 and are growing, especially after 2021. 25-44 year olds make up a growing share of the homeless population, totaling more than 50 percent of the total in 2024. At the same time, the number of homeless people in the middle-aged age group of 45.54 to 55.64 years is decreasing, and the 65+ age group is trending downwards after 2022. Children and young people (under 16) will start to decline after 2018 but will increase after 2021.

\newpage

# Discussion {#sec-data}

## Overview of Homeless in Toronto  {#sec-overview-of-Toronto-homeless}
Rising rates of homelessness in Toronto reflect significant challenges in managing housing instability, particularly among specific vulnerable groups. According to a comprehensive literature review on homelessness in Toronto, the number of homeless people has increased dramatically over the past few decades [@gaetz2004homelessness]. The data @fig-returned-to-shelter and @fig-returned-from-housing shows that non-refugees and single adults make up the largest percentage of those returning to shelters, demonstrating their vulnerability and continued housing insecurity. This trend is likely caused by economic challenges, lack of affordable housing, and inadequate social safety nets. The literature [@gaetz2004homelessness] highlights how policy changes, including cuts to social housing and benefits, have exacerbated the problem, putting more people at risk of chronic homelessness.

## Impact of unemployment on single adults aged 25-44 years {#sec-Impact-of-Unemployment}
In @fig-returned-to-shelter and @fig-returned-from-housing, we see that single adults make up a large portion of those returning to shelters for help. And from @fig-age-group-distribution-single-adult-pie and @fig-age-group-proportion-trend-yearly, 25-44 year olds make up a large portion of single adults and the overall homeless population. The unemployment crisis has severely impacted single adults, especially those aged 25-44, who make up a large portion of the homeless population.The COVID-19 pandemic triggered a sharp economic downturn that led to widespread unemployment and income insecurity. According to [@jones2020canadianlabourmarket], Canada's unemployment rate rose sharply in 2020, with young and middle-aged adults disproportionately affected. This population is particularly vulnerable because they rely heavily on employment to pay rent and meet basic needs. Without a steady income, many in this age group face eviction, bankruptcy, and an increased likelihood of homelessness. This is compounded by mental health challenges and limited access to social services [@poremski2015employment].

## Future Insights {#sec-Future-Insights}
Given the rising rates of homelessness among single adults, public policy interventions must address both immediate and long-term solutions. Expanding affordable housing and strengthening employment support programs are critical. Governments at all levels should also focus on improving access to mental health services and implementing policies that prevent evictions, especially in times of economic crises. A particular emphasis should be placed on preventative measures, such as emergency rental assistance programs and increasing the availability of affordable housing [@gaetz2004homelessness]. Additionally, the development of job retraining programs tailored to individuals who have lost employment due to economic downturns could provide sustainable solutions for those in the 25-44 age bracket.

## Weakness and Limitations {#sec-Weakness-and-Limitations}
While the data provides valuable insights into homelessness trends in Toronto, there are important limitations that should be acknowledged. The dataset reflects only individuals who use overnight shelter services, excluding those who sleep outdoors or rely on non-shelter-based services. Additionally, shelters that do not participate in Toronto's Shelter Management Information System (SMIS) are not included in the dataset. As a result, the data may underestimate the total homeless population, especially among individuals who do not seek help through formal shelter systems. 
\newpage

# Statement on LLM usage {#sec-LLMs}

Aspects of the code were written with the help of the auto-complete tool, ChatGPT. The abstract, introduction and discusstion were written with the help of ChatGPT and the entire chat history is available in other/llms/usage.txt.

# References {#sec-references}